\section*{Thank you,}

\textbf{Glenn van de Ven}, for giving me the opportunity to join you at ESO, to work again in the exciting field of dynamics, for your scientific insight, for supporting me in my work and in my scientific development and for generally believing in me. 

\textbf{Ralf Klessen}, for evaluating my thesis.

\textbf{Wilma Trick}, for your inspiring ideas for this project, for giving me scientific freedom and being patient with me, for your support in the ups and downs of this project, for providing me with many helpful advices and for believing in me when I did not. Thank you for your great supervision!

\textbf{Timo Halbesma}, for helping me with the technical set-up, issues and difficulties. 

\textbf{Laura Watkins and Prashin Jethwa}, for proofreading my thesis and for very helpful advice on my science and career.

\textbf{Volker Springel}, for providing me with the excellent Auriga simulations.

\textbf{Jo Bovy}, for providing galpy which I heavily used for my investigations.

\textbf{the European Southern Observatory and the “Galaxy Dynamics” research group in Heidelberg and Garching}, for providing such a friendly and scientifically inspiring environment and many helpful colleagues.

\textbf{the European Research Council}, for funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 724857 (Consolidator Grant ArcheoDyn).

\textbf{"ESO 5th floor"}, for making my time at ESO fun and happy.

\textbf{my parents Sigrun and Ivo}, for supporting me in everything.

\textbf{my brother and best friend Niko}, for always being there.