\documentclass[a4paper,12pt,abstracton, twoside, headings=optiontohead]{scrartcl}
%\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm]{geometry} % for printed thesis
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm, bindingoffset=15mm]{geometry}
\usepackage[onehalfspacing]{setspace}
\usepackage[ngerman, english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lipsum}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{color}
\usepackage{hyperref}
\usepackage[printonlyused]{acronym}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{empheq}
\usepackage[export]{adjustbox}
\usepackage[parfill]{parskip} 
\usepackage{pdflscape}
\usepackage{sidecap} % to have text next to figures
\sidecaptionvpos{figure}{t}
\usepackage{url}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{natbib}
\usepackage{aas_macros}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage{tabulary}

\usepackage{makecell}
\usepackage{fancyhdr} % to have the nice header and footer
%\usepackage{multirow}
\usepackage{booktabs} % nice lines in tables
\usepackage[font=small,labelfont=bf]{caption}

%#\renewcommand{\arraystretch}{1.3} 
\pagestyle{fancy}
\let\Sectionmark\sectionmark
\def\sectionmark#1{\def\Sectionname{#1}\Sectionmark{#1}}
\let\Subsectionmark\subsectionmark
\def\subsectionmark#1{\def\Subsectionname{#1}\Subsectionmark{#1}}
\let\Subsubsectionmark\subsubsectionmark
\def\subsubsectionmark#1{\def\Subsubsectionname{#1}\Subsubsectionmark{#1}}




\title{MSc thesis}
\author{Sophia Lilleengen}
\date{\today}


\addto{\captionsenglish}{\renewcommand{\bibname}{Bibliography}}


\begin{document}

\bibliographystyle{glenn_new} 

\defcitealias{Bland-Hawthorn...MW...2016}{B-H\&G16}
\defcitealias{AurigaGrand}{G+17}
\defcitealias{Vanhollebeke...bulge...2009}{VGG09}
\defcitealias{Gibbons...sagstream...2014}{Gibbons+ '14}
\defcitealias{Dierickx...sagstream..2017}{Dierickx+ '17}
\defcitealias{Newberg...orphanstream..2010}{Newberg+ '10}
\defcitealias{Koposov...GD1stream...2010}{Koposov+ '10}
\defcitealias{Bowden...GD1stream...2015}{Bowden+ '15}
\defcitealias{Malhan...GD1stream...2018}{Malhan+ '18}
\defcitealias{Kupper...pal5stream...2015}{K\"upper+ 15}
\defcitealias{Streams..GD1..Pal5...Bovy...2016}{Bovy+ '16}
\defcitealias{MWmass...GCmotions...Watkins...2018}{Watkins+ '18}
\defcitealias{Posti...MWmassGCs...2019}{Posti+ '19}
\defcitealias{Vasiliev...GCoverview...2018}{Vasiliev+ '18}
\defcitealias{Sohn...GCsHST..2018}{Sohn+ '18}
\defcitealias{GaiaDR...GCs...2018}{Gaia+ '18b}
\defcitealias{MWmass...sat...dyn}{Callingham+ '18}
\defcitealias{Eadie...GCsBayes...2018}{Eadie+ '18}


\pagenumbering{roman}

\include{titlepages-eng}

\newpage
\thispagestyle{empty}
\begin{abstract}
\hspace{-12pt}Many astrophysical and galaxy-scale cosmological problems require a well-determined gravitational potential. Globular clusters (GCs) surrounding \linebreak galaxies can be used as dynamical tracers of the luminous and dark matter distribution at large (kpc) scales. This M.Sc. project investigates - by means of the Auriga galaxy simulations and in anticipation of high-resolution IFU data of external galaxies - whether a novel action-based approach could provide a constraint for an axisymmetric approximation of the gravitational potential. \\
In an axisymmetric potential, actions (radial $J_R$, vertical $J_z$ and angular momentum $L_z$) are integrals of motion and can be used to characterize and label orbits. In the Milky Way (MW), the assumption that stars in cold streams are on similar orbits was found to be a useful first-order constraint of its gravitational potential. In external galaxies, no individual stars but only GCs can be resolved. One could expect GCs from the same dwarf galaxy (DG) merger event to move at the present time on similar orbits in the host galaxy, analogously to stellar streams in the MW, and should therefore have similar actions in the true (axisymmetric) potential. \\
We investigate this idea in one galaxy of the cosmological N-body simulation suite Auriga \citep{AurigaGrand}. As a first step, we present an effective strategy to fit analytic, axisymmetric, time-dependent potential models with slowly varying parameters to the simulation that are good enough to estimate actions. \\
Then, we select stellar particles born in dwarf galaxies as proxies for GCs and follow the evolution of their orbital actions during the process of merging with a more massive galaxy. These actions show a significant variation over time. As a result, at $z=0$, the stellar particles accreted in the same merger event show a very extended distribution in action space. We find that minimizing this distribution, however, cannot constrain the true potential since actions and their evolution are affected by complex physical processes during mergers. \\
In local observations, we confirm this result in the stars of Gaia-Enceladus, one of the few DG mergers of our MW that we know of. Their action distribution is smeared out extensively. Based on these results, we propose that modellers need to find and develop more realistic distribution functions for GCs of a single DG merger event in simulations before being able to constrain the gravitational potential of external galaxies using action-based dynamical modelling of GCs.

\newpage
%\vspace{2cm}
\thispagestyle{empty}
\begin{center}
 \textbf{Zusammenfassung}\end{center}

\hspace{-12pt}Viele astrophysikalische und kosmologische Probleme auf Ga-laxien Skalen erfordern ein genaues Gravitationspotential. Kugelsternhaufen (GCs), die Galaxien umgeben, k\"onnen als dynamische Tracer für die Verteilung von Licht und dunkler Materie in großen (kpc) Skalen verwendet werden. Diese Masterarbeit untersucht in der Galaxiensimulation Auriga und in Erwartung hoch-aufgel\"oster IFU-Daten externer Galaxien, ob ein neuartiger wirkungsbasierter Ansatz eine Einschr\"ankung für eine axialsymmetrische Approximation des Gravitationspotenzials darstellen k\"onnte. \\
In einem achsensymmetrischen Potential sind Wirkungen (radiale $ J_R $, vertikale $ J_z $ und der Drehimpuls $ L_z $) Bewegungsintegrale und k\"onnen verwendet werden, um Bahnen zu charakterisieren und zu kennzeichnen. In der Milchstraße (MW) wurde die Annahme, dass Sterne in kalten Str\"omen auf \"ahnlichen Bahnen liegen, als n\"utzliche Einschränkung erster Ordnung für ihr Gravitationspotenzial befunden. In externen Galaxien k\"onnen keine einzelnen Sterne, sondern nur GCs aufgel\"ost werden. Man k\"onnte erwarten, dass sich GCs aus demselben Zwerggalaxie(DG)-Fusionsereignis auf \"ahnlichen Bahnen in der Wirtsgalaxie bewegen, analog zu Sternstr\"omen in der MW, und sollten daher \"ahnliche Wirkungen auf das wahre (axialsymmetrische) Potenzial haben. \\
Wir untersuchen diese Idee in einer Galaxie der kosmologischen N-K\"orper-simulationsserie Auriga \citep{AurigaGrand}. Im ersten Schritt pr\"asentieren wir eine effektive Strategie, um analytische, achsensymmetrische, zeitabh\"angige Potenzialmodelle mit langsam variierenden Parametern an die Simulation an-zupassen, die für die Absch\"atzung von Wirkungen geeignet sind. \\
Dann w\"ahlen wir Sternpartikel, die in Zwerggalaxien geboren sind, als Stellvertreter für GCs und verfolgen die Entwicklung ihrer Umlaufbewegungen w\"ahrend der Verschmelzung mit einer massereicheren Galaxie. Diese Wirkungen zeigen im Laufe der Zeit einen erheblichen Ver\"anderung. Bei $ z = 0 $ zeigen die im selben Fusionsereignis angeh\"auften Sternpartikel daher eine sehr ausgedehnte Verteilung im Wirkungsraum. Wir haben festgestellt, dass das Minimieren dieser Verteilung das wahre Potenzial jedoch nicht einschr\"anken kann, da Wirkungen und deren Entwicklungen von komplexen physikalischen Prozessen w\"ahrend der Verschmelzung beeinflusst werden. \\
In lokalen Beobachtungen best\"atigen wir dieses Ergebnis in den Sternen von Gaia-Enceladus, einer der wenigen Zwerggalaxienfusionen unserer MW, die wir gut kennen. Ihre Wirkungsverteilung ist weitgehend verschmiert. Basierend auf diesen Ergebnissen schlagen wir vor, dass Modellierer in Simulationen realistischere Verteilungsfunktionen für GCs eines einzelnen DG-Fusionsereignisses suchen und entwickeln müssen, bevor sie das Gravitationspotenzial externer Galaxien mithilfe wirkungsbasierter dynamischer Modellierung von GCs einschränken können.

\end{abstract}
\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\newpage
\fancyhf{}
\fancyhead{}
\fancyhead[RO,LE]{Contents}
\fancyfoot{}
%depends on how many pages before the stars are, 1 could start on an even page what it does here.......
\fancyfoot[LO,RE]{}
\fancyfoot[LE,RO]{\thepage}

\addtocontents{toc}{}
\tableofcontents
\newpage
\pagenumbering{arabic}

\fancyhf{}
\fancyhead{}
\fancyhead[LE]{\thesubsection.\ \Subsectionname}
\fancyhead[RO]{\thesection.\ \Sectionname}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,RE]{}
\include{Introduction}


\include{Auriga}

\include{Dynamics}

\include{Discussion}

\fancyhf{}
\fancyhead{}
\fancyhead[RO,LE]{5. Summary \& conclusion}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,RE]{}
\include{Sum_Conc}

\newpage
\fancyhf{}
\fancyhead{}
\fancyhead[RO,LE]{Acronyms}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,RE]{}
\section*{Acronyms}
\addcontentsline{toc}{section}{Acronyms}

\begin{acronym}[WIMPs]
    \acro{AGN}{active galactic nuclei}
    \acro{AMR}{age-metallicity relation}
    \acro{CBE}{collisionless Boltzmann equation}
    \acro{CDM}{cold dark matter}
    \acro{CMB}{cosmic microwave background}
    \acro{DF}{distribution function}
    \acro{DG}{dwarf galaxy}
	\acro{DM}{dark matter}
	\acro{DMO}{dark matter only}
	\acro{D/T}{disk-to-total}
	\acro{GC}{globular cluster}
	\acro{HST}{Hubble Space Telescope}
	\acro{IFU}{Integral Field Unit}
	\acro{IoM}[IoMs]{integrals of motion}
	\acro{LCDM}[$\Lambda$CDM]{Lambda cold dark matter}
	\acro{LG}{Local Group}
	\acro{MGE}{Multi-Gaussian Expansion}
	\acro{MHD}{magnetohydrodynamical}
	\acro{ML}[M/L]{mass-to-light}
	\acro{MN}{Miyamoto-Nagai}
	\acro{MoND}{Modified Newtonian Dynamics}
	\acro{MW}{Milky Way}
	\acro{NFW}{Navarro-Frenk-White}
	\acro{PM}{proper motion}
	\acro{TF}{Tully-Fisher}
	\acro{WDM}{warm dark matter}
	\acro{WIMPs}{Weakly Interacting Massive Particles}

\end{acronym}
\newpage

\fancyhf{}
\fancyhead{}
\fancyhead[RO,LE]{References}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,RE]{}
\addcontentsline{toc}{section}{References}

\bibliography{mybib}{}
%\bibliographystyle{mnras}
\newpage
\addcontentsline{toc}{section}{Acknowledgements}
\thispagestyle{plain}
\include{acknowledgements}
\newpage
\addcontentsline{toc}{section}{Statement}
\thispagestyle{plain}
\include{deposition}


\end{document}
